SMT-Based Generation of Symbolic Automata \thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-03-30 2017-Xudong-SMT/ActaInformatica.tex
Acta Inf. paper (minor): moved the hyperref package to avoid name use errors at compilations
Improve the proof of Th.3 after Skype.

%\section{}
%\subsection{}

@@

\paragraph{States:} for each of these, define $States: \langle OT\rangle\rightarrow\langle State\rangle$ which is the function computing the set of states in a
set of OTs.

@@

Define the function $States$ as before, and  $Assignments : \langle OT\rangle\times\langle State\rangle\rightarrow\langle Assign\rangle$ the function
computing the set of assignments of a global state from a given set of OTs.

The structure implicitely computed by the algorithm is a triple $ Aut= \langle OTs, States, Assigns\rangle$.

Then the assignments set of $OA_1$ is $Assignments(L1, States(L1))$ and that of $OA_2$ is $Assignments(L2_{sat}, States(L2_{sat}))$. The computed triples are 
\[Aut_1 = \langle L1_{sat}, States(L1_{sat}), Assignments(L1, States(L1))\rangle\] and \[Aut_2 = \langle L2_{sat}, States(L2_{sat}), Assignments(L2_{sat}, States(L2_{sat}))\rangle \ . \]

Based on the notation of the triple, we define the function $Refine : Aut \rightarrow Aut$ refining the result automaton together with the assignments,
\begin{align*}
Refine(Aut) &=\langle OT', States(OT'), Assignments(OT', States(OT'))\rangle \\
&=Aut'
\end{align*}
where $OT' = OT - OT_{unsat}$, $OT_{unsat}$ is the set of transitions checked unsatisfiable in this iteration.

Given that every iteration will remove some unsatisfiable transitions, $Refine$ is a monotone function. There exists a sequence $Aut^n = Refine(Aut^{n-1})$ s.t. $|Aut^n| < |Aut^{n-1}|$.
Let $Aut''$ contains the $OT''$ which has no more unsatisfiable transition, then there is $Aut'' = Refine(Aut'')$ and $|Aut''| = |Refine(Aut'')|$.
Let $Aut''_1$ and $Aut''_2$ be the two
 fixed points with $Refine(Aut''_1) = Aut''_1$, $Refine(Aut''_2) = Aut''_2$.

For the transition sets in $Aut''_1$ and $Aut''_2$, $L1''_{sat}=L1_{reach}$ and $L2''_{sat}=L2_{reach}$, we need to check their equivalence.
At the first iteration, the unsatisfiable transitions in $OA_1$ has only one case according to the relation between the assertions to check,
\begin{align*}
&\quad Pred\wedge \bigvee_i Assignments_i(L1, States(L1) \text{ is false}\\
&\quad\longrightarrow Pred\wedge \bigvee_j Assignments_j(L2_{sat}, States(L2_{sat})\text{ is false}\\
&\Rightarrow \forall t\in L1_{unsat} \longrightarrow t\in L2_{unsat}\\
&\Rightarrow L1_{unsat}\subseteq L2_{unsat}.
\end{align*}
However, its reverse direction is not fulfilled. Recall that $L1_{sat}=L2_{sat}$, these relations can imply that,
\begin{align*}
Assignments(L1, &States(L1))\supseteq Assignments(L2_{sat}, States(L2_{sat}))\\
&\supseteq Assignments(L1_{sat}-L1_{unsat}, States(L1_{sat}-L1_{unsat}))\\
&\supseteq Assignments(L2_{sat}-L2_{unsat}, States(L2_{sat}-L2_{unsat}))
\end{align*}
So after the first iteration, the unsatisfiable transitions in $OA_1$ has two cases,
\begin{align*}
&\forall t\in L1'_{unsat} \longrightarrow t\in L2'_{unsat}\\
\text{ or }&\forall t\in L1'_{unsat} \longrightarrow t\notin L2'_{sat}.
\end{align*}
By induction, after removing all the transitions not in $L2''_{sat}$, we obtain  $L1_{reach}=L2_{reach}$, which implies that $States_{L1_{reach}}=States_{L2_{reach}}$ and
$$Assignments\\(L1_{reach},States_{L1_{reach}})=Assignments(L2_{reach},States_{L2_{reach}})\ .$$
\hfill\qed
\end{proof}

@@

  \medskip
  \noteInEM{I keep below the original proof draft from Xudong, though
    think it does not apply any more}

@@

\begin{figure}[ht]
\centerline{ 
  \includegraphics[width=0.30\linewidth]{ActaXFIG/Enable1}
  \hspace{25mm}
  \includegraphics[width=0.25\linewidth]{ActaXFIG/Enable2}}
  \begin{minipage}{5.7cm}
$  \begin{array}{@{}r@{}l@{}}
    ot_1  = &\openrule{\{ \xrightarrow{a1}_P \},\\ [\forall x1. a1\neq\delta(x1)]}
    {A_0 \xrightarrow{a1} A_0}\\
    ot_2  =& \openrule{
                            \{\xrightarrow{\delta(x2)}_P,\,
                            \xrightarrow{acc(x2)}_Q\}  }
    {A_0 \xrightarrow{\underline{\delta(x2)}} A_1}\\
    ot_3  =& \openrule{
                          \{  \xrightarrow{a3}_q \},\\ [\forall x3. a3\neq\delta(x3)] }
    {A_1 \xrightarrow{a3} A_1}
  \end{array}$
  \end{minipage}
  \hspace{2mm}
  \begin{minipage}{6cm}
$  \begin{array}{@{}r@{}l@{}}
    ot'_1  =& \openrule{
    	\{\xrightarrow{b1}_P\}, ~~ [\forall y1. b1\neq\delta(y1) \!\land\! v=0]}
    {B_0 \xrightarrow{b1} B_0}\\
    ot'_2  =& \openrule{
                            \{\xrightarrow{\delta(y2)}_P,\,
                            \xrightarrow{acc(y2)}_Q \}, \\
                                        v=0, \\
                                        \{v\gets 1\}}
    {B_0 \xrightarrow{\underline{\delta(y2)}} B_1}\\
    ot'_3  =& \openrule{
                            \xrightarrow{b2}_q, \\ [\forall y2. b2\neq\delta(y2) \!\land\! v=1] }
    {B_1 \xrightarrow{b2} B_1}
  \end{array}$
  \end{minipage}
  \caption{The two open automata}
  \label{schema:enable-OAs}
\end{figure}

>>>>

%%\section{}
%%\subsection{}

@@

\paragraph{States:} for each of these, define $States: \langle OT\rangle\rightarrow\langle State\rangle$ which is the function computing the set of target states in a
set of OTs.

@@

Define the function $States$ as before, and  $Assignments : \langle OT\rangle\times State\rightarrow\langle Assign\rangle$ the function
computing the set of assignments of a global state from a given set of OTs.

The structure implicitely computed by the algorithm is a triple $ Aut= \langle OTs, States,\\ Assigns\rangle$.

Then the assignments set of $OA_1$ is $\{Assignments(L1, State_{i})\}^{State_{i}\in States(L1)}$ and that of $OA_2$ is $\{Assignments(L2_{sat}, State_{j})\}^{State_{j}\in States(L2_{sat})}$. The computed triples are 
\[Aut_1 = \langle L1_{sat}, States(L1_{sat}), \{Assignments(L1, State_{i})\}^{State_{i}\in States(L1)}\rangle\] and \[Aut_2 = \langle L2_{sat}, States(L2_{sat}), \{Assignments(L2_{sat}, State_{j})\}^{State_{j}\in States(L2_{sat})}\rangle \ . \]

Based on the notation of the triple, we define the function $Refine : Aut \rightarrow Aut$ implementing $makeReachableSubAutomaton(S0,OTs)$, it uses the assignments of the previous step to refine the computed reachable space.
\begin{align*}
Refine(Aut) &=\langle L', States(L'), \{Assignments(L', State_{i})\}^{State_{i}\in States(L')}\rangle \\
&=Aut'
\end{align*}
where $L' = L - L_{unsat}$, $L_{unsat}$ is the set of transitions checked unsatisfiable in this iteration. So there must has $|L'|<|L|$, and potentially, their target states and the corresponding assignments may be reduced: 
\[States_{L'}=\{State_{init}\}\cup \bigcup States(L'),\]
\[Assignments_{L',States_{L'}}=\bigcup\{Assignments(L', State_{i})\}^{State_{i}\in States_{L'}} \]
\[s.t. |States_{L'}|\le|States_{L}|\wedge|Assignments_{L',States_{L'}}|\le|Assignments_{L,States_{L}}|.\]

In one iteration, satisfiability of every transition $\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\in L$ is checked through judging whether 
\[\bigvee_{Assign_i\in Assignments_{L, States_{L}}}(Pred_{l}\wedge Assign_{i})\text{ is true}.\]
More transitions may be checked unsatisfiable with the reduction of the assignments set, so the function need iterate several times. Given that in every iteration, the number of triples decreases, $Refine$ is a monotone function. And there exists a sequence \[Aut^n = Refine(Aut^{n-1})\quad s.t. |Aut^n| < |Aut^{n-1}|.\]
The function terminates if $Aut''$ contains the no more unsatisfiable transition, then there is the fixed-point $Aut'' = Refine(Aut'')$ and $|Aut''| = |Refine(Aut'')|$.

Let $Aut''_1$ and $Aut''_2$ be the two fixed-points with $Refine(Aut''_1) = Aut''_1$, $Refine(Aut''_2) = Aut''_2$.For the transition sets in $Aut''_1$ and $Aut''_2$, $L1''_{sat}=L1_{reach}$ and $L2''_{sat}=L2_{reach}$, we need to check their equivalence.

First, there is an inclusion relation between the initial assignments sets, 
\begin{align*}
&L2_{sat}\subseteq L1\Rightarrow States_{L2_{sat}}\subseteq State_{L1}\\
\Rightarrow &Assignments_{L2_{sat}, States_{L2_{sat}}}\subseteq Assignments_{L1, States_{L1}}
\end{align*}

Then we see what changes have occurred during the iterations. In the first iteration, the inclusion relation between their assignments gives the the inclusion relation of unsatisfiable transitions, for $\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}} \in L1_{unsat}$ and $\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}} \in L2_{unsat},$
\begin{align*}
&\bigvee_{Assign_i\in Assignments_{L1, States_{L1}}} (Pred_{l}\wedge Assign_i)\text{ is false}\\
\longrightarrow&\bigvee_{Assign_j\in Assignments_{L2_{sat}, States_{L2_{sat}}}} (Pred_{l}\wedge Assign_j)\text{ is false}.
\end{align*}
The relation that $\forall\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}} \in L1_{unsat},\exists\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}} \in L2_{unsat}$, implies that $L1_{unsat}\subseteq L2_{unsat}.$

However, its reverse direction is not fulfilled. Recall that $L1_{sat}=L2_{sat}$, more inclusion relations of the assignments can be implied,
\begin{align*}
Assignments_{L1, States_{L1}}&\supseteq Assignments_{L2_{sat}, States_{L2_{sat}}}\\
&\supseteq Assignments_{L1_{sat}-L1_{unsat}, States_{(L1_{sat}-L1_{unsat})}}\\
&\supseteq Assignments_{L2_{sat}-L2_{unsat}, States_{(L2_{sat}-L2_{unsat})}}
\end{align*}

Let $L1'_{sat}=L1_{sat}-L1_{unsat}$ and $L2'_{sat}=L2_{sat}-L2_{unsat}$
Then in the next iteration, the unsatisfiable transition $\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}$ removed from $L1'_{sat}$ might be the same transition (up to renaming) removed from $L2'_{sat}$, that is
\[\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\in L1'_{unsat} \longrightarrow \ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\in L2'_{unsat}.\]
Or it is the transition has not removed in the last iteration because of 
\[Assignments_{L2_{sat}, States_{L2_{sat}}}\subseteq Assignments_{L1, States_{L1}}.\] 
We have already proved that 
\[Assignments_{L1'_{sat}, States_{L1'_{sat}}}\subseteq Assignments_{L2_{sat}, States_{L2_{sat}}},\]
so those transitions will be removed this time, that is
\[\ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\in L1'_{unsat} \longrightarrow \ostate{\tilde{t}}\xrightarrow{l}\ostate{\tilde{t'}}\notin L2_{sat}.\]

By induction, if there is no more unsatiable transitions in $L2''_{sat}$, $L1''_{sat}$ continues removing all the transitions not in $L2''_{sat}$, until $L1_{reach}=L2_{reach}$, which implies that $States_{L1_{reach}}=States_{L2_{reach}}$ and
$$Assignments_{L1_{reach},States_{L1_{reach}}}=Assignments_{L2_{reach},States_{L2_{reach}}}.$$
\hfill\qed
\end{proof}

@@

%  \medskip
%  \noteInEM{I keep below the original proof draft from Xudong, though
%    think it does not apply any more}

@@

\begin{figure}[ht]
\centerline{ 
  \includegraphics[width=0.30\linewidth]{ActaXFIG/Enable1}
  \hspace{25mm}
  \includegraphics[width=0.25\linewidth]{ActaXFIG/Enable2}}
  \begin{minipage}{5.7cm}
$  \begin{array}{@{}r@{}l@{}}
    ot_1  = &\openrule{\{ \xrightarrow{a1}_P \},\\ [\forall x1. a1\neq\delta(x1)]}
    {A_0 \xrightarrow{a1} A_0}\\
    ot_2  =& \openrule{
                            \{\xrightarrow{\delta(x2)}_P,\,
                            \xrightarrow{acc(x2)}_Q\}  }
    {A_0 \xrightarrow{\underline{\delta(x2)}} A_1}\\
    ot_3  =& \openrule{
                          \{  \xrightarrow{a3}_q \},\\ [\forall x3. a3\neq\delta(x3)] }
    {A_1 \xrightarrow{a3} A_1}
  \end{array}$
  \end{minipage}
  \hspace{2mm}
  \begin{minipage}{6cm}
$  \begin{array}{@{}r@{}l@{}}
    ot'_1  =& \openrule{
    	\{\xrightarrow{b1}_P\}, ~~ [\forall y1. b1\neq\delta(y1) \!\land\! v=0]}
    {B_0 \xrightarrow{b1} B_0}\\
    ot'_2  =& \openrule{
                            \{\xrightarrow{\delta(y2)}_P,\,
                            \xrightarrow{acc(y2)}_Q \}, \\
                                        v=0, \\
                                        \{v\gets 1\}}
    {B_0 \xrightarrow{\underline{\delta(y2)}} B_1}\\
    ot'_3  =& \openrule{
                            \xrightarrow{b2}_q, \\ [\forall y2. b2\neq\delta(y2) \!\land\! v=1] }
    {B_1 \xrightarrow{b2} B_1}
  \end{array}$
  \end{minipage}
  \caption{The two open automata}
  %% \noteSB{I guess pNets in this figure have to be replaced by open
  %%   automata}
  \label{schema:enable-OAs}
\end{figure}
